Use and acquisition of causal knowledge in a decision making process

### 2018-11-27 Propuesta/Presentacion/Presentacion_Corta.tex
ultimas correcciones

}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
%\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage{natbib}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposition}
\newtheorem{defi}[teo]{Definition}
\newtheorem{obs}[teo]{Observation}
\newtheorem{lem}[teo]{Lemma}
\newtheorem{cor}[teo]{Corolary}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}

@@

\title[Short title]{Use and acquisition of causal knowledge in a decision making process} % The short title appears at the bottom of every slide, the full title is only on the title page

@@

\begin{frame}
\frametitle{Human use of Causal Relations}
\begin{itemize}
\item \cite{hagmayer2013repeated} show how human beings use and modify causal information.
\item Both \cite{hohwy2013predictive}, \cite{clark2015surfing}, assert that the brain itself is a causal hypothesis tester. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our setting}
\begin{itemize}
\item Consider a \textit{rational} decision maker in an uncertain environment.
\item Consequences of actions can not be predicted.
\item We add an extra to the environment: causal structure. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A brief on causal models}
\begin{itemize}
\item Causal models encode cause-effect relations. 
\item Relations are encoded in a graph.
\item Can be used to predict \textit{effects} of certain causes.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Our Proposal}
\begin{itemize}
\item Current decision-making theories do not consider causal structure in the decision making processes
\item We \textbf{propose} an autonomous rational agent who can use and learn causal information while interacting with its environment.
\end{itemize}
\end{frame}

\section{Concepts}
\begin{frame}
\frametitle{Concepts}
Concepts and Framework:
\begin{enumerate}
\item Causal Models
\item Decision Problems
\end{enumerate}
\end{frame}

\subsection{Causal Models}
		\begin{frame}
		\frametitle{Difference between causal and associative relations}
		\begin{itemize}
		\item With associative relations we can't figure out what influenced (or caused) what.
		\item Consider this: A distribution $p(a,t)$ can be factored both as $p(a|t)p(t)$ and $p(t|a)p(a)$. 
		\item In terms of a Bayesian Network, $T \to A$, and $A \to T$, but both can't be causal.
		\item Causal relations need not be in temporal order: consider a barometer and a storm.
		\end{itemize}
		\end{frame}

		\begin{frame}
		\frametitle{Formal Definition of Causality according to \cite{spirtes2000causation}}
		\begin{defi}
		Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite, probability space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
		\begin{itemize}
		\item Transitive: if $A \to B$ and $B \to C$ for $A,B,C \in \mathcal{F}$ then $A \to C$.
		\item  Irefflexive: for each $A \in \mathcal{F}$ it is not true that $A \to A$.
		\item Antisymetric: For $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
		\end{itemize}
		If $A \to B$ we will say that $A$ is a cause of $B$.
		\end{defi}
		\end{frame}

		\begin{frame}
		\frametitle{From Graphs to Probability Measures}
		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph. We require that this measure satisfies:
        \begin{itemize}
        \item Causal Markov Condition.
        \item Causal Minimality.
         \item Causal Faithfulness.
         \item Causal Sufficiency
        \end{itemize}
         This requirements will be treated in an axiomatic fashion.
		\end{frame}

@@

\subsection{Decision Problems}
	\begin{frame}
	\frametitle{Decision Problems}
	A decision problem is composed by (\cite{bernardo2000bayesian}):
	\begin{itemize}
	\item A set of available options $\mathcal{A}$.
	\item A (rational) decision maker with preference relation $\succeq$ defined over $\mathcal{A} \times \mathcal{A}$.
	\item A family of uncertain events $\mathcal{E}$.
    \item A family of possible outcomes $\mathcal{C}$
	\end{itemize}
	\end{frame}

@@

\section{Research Problem}
\begin{frame}
About my Proposed Research
\end{frame}
	\subsection{Motivation}
	\begin{frame}
	\frametitle{Motivation: 1}
	\begin{itemize}
	\item Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}).
	\item Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions.
	\end{itemize}
	\end{frame}
	\begin{frame}
	\frametitle{Motivation: 2}
	\begin{itemize}
	\item It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
	\item What about not knowing the causal model? 
	\end{itemize}
	\end{frame}

	\subsection{Justification}
	\begin{frame}
	\frametitle{Justification: Un-interpretable models}
	\begin{itemize}
	\item Many real-world applications of decision making are solved by \textit{associative} methods.
	\item Associative methods capture only statistical patterns that are found in data.
	\item Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
	\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Justification: Need for nterpretable decision-making models}
	\begin{itemize}
	\item Current methods for Decision Making are based in Reinforcement Learning, and specifically in Deep Reinforcement Learning. 
	\item Although good performance in the task, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
	\item This is highly relevant in real-world applications such as self-driving cars.
	\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Justification: Explaining why}
	\begin{itemize}
	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain. 
	\item It allows to ask \textit{why}.
	\item Transferability
	\end{itemize}
	\end{frame}

@@

\section{Approaches and Related Work}
\begin{frame}
\frametitle{Approaches and Related Work}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Related_Work.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame} 
\subsection{Approaches and Related Work: Reinforcement Learning}
\begin{frame}
\frametitle{Reinforcement Learning (RL)}
\begin{itemize}
\item Reinforcement Learning uses only associative patterns for finding an optimal policies.
\item Optimal policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
\item In this way, when only associative information is available, RL is a coherent learning framework. 
\end{itemize}
\end{frame}

\subsection{Approaches and Related Work: Causal learning in humans}
\begin{frame}
\frametitle{Causal Learning in Human beings}
\begin{itemize}
\item Human beings, while facing a sequential decision problem, use and modify causal information.
\item \cite{hagmayer2009decision}, \cite{hagmayer2013repeated} show that human beings can learn causal information in simple decision-making tasks, but remains open to study causal learning while seeking for a specific objective. 
\item 
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Joyce's Causal Decision Theory}
\begin{frame}
\frametitle{Joyce's Causal Decision Theory}
\begin{itemize}
\item An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes.
\item  His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Bandits}
\begin{frame}
\frametitle{Bandits: Non-causal}
\begin{itemize}
\item In the class of decision problems considered for this proposal it is assumed that the a rational agent must choose some available action and then receive a reward from it.
\item Problems of this nature have been modelled as \textit{bandit} problems.
\item Best-arm algorithms exist: none consider causal information.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Bandits: Causal bandits}
\begin{itemize}
\item An action can be conceived as an \textit{intervention} over an environment, which is in fact how humans consider their actions in the world.
\item They use and modify causal knowledge during a sequential decision making process.
\item It is shown by \cite{lattimoreNIPS2016} that adding causal information in a fixed budget decision problem allows the decision maker to learn \textit{faster} had he not considered causal information.
\item Their work requires that the causal model is fully known to the decision maker.
\item this requirement is relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part. 
\end{itemize}
\end{frame}

\section{Proposed Solution and Methodology}
\begin{frame}
\frametitle{Proposed Solution and Methodology}
\end{frame}
	\subsection{Objectives}
	\begin{frame}
	\frametitle{General Objectives}
	The proposed research has as a general objective:
	\begin{itemize}
 \item Understand the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
	\end{itemize}
	\end{frame}

	\subsection{Specific Objectives}
	\begin{frame}
	\frametitle{Specific Objectives}
	In order to reach the General Objective we must:
	\begin{enumerate}
	\item Game between agent and Nature in order to modell agent-environment interaction.
	\item Find an adequate family of probability distributions that encode causal knowledge.
	\item General belief-updating criteria in terms of causal information.
	\end{enumerate}
	\end{frame}

@@

\begin{frame}
\frametitle{A conjecture}
It needs to be shown that given the rewards obtained by a decision-making procedure that achieves the MEU, then my proposed algorithm must stay close to it.
\begin{teo}
Let $(X_1,X_2,...)$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...)$ are the rewards obtained by a decision-making procedure that uses causal information, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$. 
\end{teo}
\begin{proof}
To be proved.
\end{proof}
\end{frame}

@@

\subsection{Proposed Solutuion: A guiding principle}
\begin{frame}
\frametitle{A guiding principle for Causal Decision Problems}
We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational. We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable. It remains to be refined and extended for a general case. 
\end{frame}

\begin{frame}
\begin{prop}
Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent. Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
\end{prop}
\end{frame}

\subsection{First Step: Modelling a Causal Decision Problem as a Game}
\begin{frame}
\frametitle{Games}
\begin{itemize}
\item In this way, the first step in order to solve the stated problem is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
\begin{enumerate}
\item the (original) decision maker, 
\item and a player which will be called Nature, whose actions are to be guided by the causal model.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item  Since in decision problems under uncertainty it is assumed that the \textit{state} of the environment is unknown, we will assume that player Nature has the first move and he assigns some state to any of the variable in the causal model $\mathcal{G}$.
\item After the decision maker makes his \textit{play}, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
\item This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games
\end{itemize}
\end{frame}

\subsection{Second Step: Belief formation}
\begin{frame}
\frametitle{Belief Formation}
\begin{itemize}
\item The whole idea is that a decision maker is able to learn from his environment by observing, and reasoning about, the effects of his actions.
\item The observed information will modify his current knowledge, which is encoded as probabilistic beliefs about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world
\end{itemize}
\end{frame}

\subsection{Third Step: Making choices}
\begin{frame}
\frametitle{Making Choices}
\begin{itemize}
\item The decision maker will use his current knowledge \textit{as if} it were the truth about the causal relations of his environment.
\item With this information he will choose the action that maximize the Expected Utility, according with our guiding principle.
\end{itemize}
\end{frame}

\subsection{Third Step: Belief updating}
\begin{frame}
\frametitle{Belief Updating}
\begin{itemize}
\item If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms.
\item it remains to be answered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way
\end{itemize}
\end{frame}

\section{Methodology divided into simpler cases}
\begin{frame}
We will show how the methodology applies in simpler and easier cases where the decision maker has some exta info about his environment.
\end{frame}
\subsection{Fully knowing the causal graphical model}
\begin{frame}
\frametitle{Fully knowing the causal graphical model}
\begin{itemize}
\item If the decision maker knows the causal graphical model, then following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
\end{itemize}
\end{frame}

\subsection{Knowing only the graph structure}
\begin{frame}
\frametitle{Knowing only the causal graph structure}
\begin{itemize}
\item If a decision maker knows the structure of the graph, but he does not know the parameters,
\item then he encodes this ignorance, along with any useful previous knowledge, as a probability distribution over a suitable space.
\item In general, what must be specified is a way of encoding current causal knowledge into a probability distribution in such a way that a \textit{local} (i.e. in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
\item This means that using current beliefs, the agent forms a causal model which is used as if it were the true causal model.
\end{itemize}
\end{frame}

\subsection{General case: Model unknown}
\begin{frame}
\frametitle{Model unknown}
If the model is completely unknown, two ideas appear:
\begin{enumerate}
\item If the max number of variables is known.
\item If it is not.
\end{enumerate}
\end{frame}
\begin{frame}
\frametitle{Idea 1 for the General Case}
\begin{itemize}
\item If the maximum number of variables is assumed to be known by the decision maker, then the situation simplifies since a decision maker could generate distributions that represent causal relations among variables.
\item For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model as in the previous case.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Idea 2 for the General Case}
\begin{itemize}
\item If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables. In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.  
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{20 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/20_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

>>>>

}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
%\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage{natbib}
\theoremstyle{plain}
\newtheorem{teo}{Theorem}
\newtheorem{prop}[teo]{Proposition}
\newtheorem{defi}[teo]{Definition}
\newtheorem{obs}[teo]{Observation}
\newtheorem{lem}[teo]{Lemma}
\newtheorem{cor}[teo]{Corolary}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}

@@

\title[Causally Choosing]{Use and acquisition of causal knowledge in a decision making process} % The short title appears at the bottom of every slide, the full title is only on the title page

@@

\begin{frame}
\frametitle{Human use of Causal Relations}
\begin{itemize}
\item \cite{hagmayer2013repeated} show how human beings use and modify causal information.
\item Both \cite{hohwy2013predictive}, \cite{clark2015surfing}, assert that the brain itself is a causal hypothesis tester. 
\item While Danks \cite{danks2014unifying} argues that cognitive processes (including causal reasoning) take place in the mind as graphical models, Tenenbaum's school argues that the mind uses hierarchical bayesian inference (\cite{kemp2010learning}).
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our setting}
\begin{itemize}
\item Consider a \textit{rational} decision maker in an uncertain environment.
\item Consequences of actions can not be predicted.
\item We add an extra to the environment: causal structure.
\item This means that there exists a fix but unknown causal model that governs the environment where the agent lives.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A brief on causal models}
\begin{itemize}
\item Causal models encode cause-effect relations. 
\item Relations are encoded in a graph.
\item Can be used to predict \textit{effects} of certain causes without necessarily performing the action.
\item They allow to \textit{imagine} what would happen if...
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Our Proposal}
\begin{itemize}
\item Current decision-making theories do not consider causal structure in decision making processes
\item We \textbf{propose} an autonomous rational agent who can use and learn causal information while interacting with its environment.
\end{itemize}
\end{frame}

\section{Concepts}
\begin{frame}
\frametitle{Concepts}
Concepts and Framework:
\begin{enumerate}
\item Causal Models.
\item Decision Problems.
\end{enumerate}
\end{frame}

\subsection{Causal Models}
		\begin{frame}
		\frametitle{Causal Models}
		Causality and Causal Models:
		\end{frame}
		\begin{frame}
		\frametitle{Difference between causal and associative relations}
		\begin{itemize}
		\item With associative relations we can't figure out what influenced (or caused) what.
		\item Consider this: A distribution $p(a,t)$ can be factored both as $p(a|t)p(t)$ and $p(t|a)p(a)$. 
		\item In terms of a Bayesian Network, $T \to A$, and $A \to T$, but both can't be causal.
		\item Causal relations need not be in temporal order: consider a barometer and a storm.
		\end{itemize}
		\end{frame}

		\begin{frame}
		\frametitle{Formal Definition of Causality according to \cite{spirtes2000causation}}
		\begin{defi}
		Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite, probability space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
		\begin{itemize}
		\item Transitive: if $A \to B$ and $B \to C$ for $A,B,C \in \mathcal{F}$ then $A \to C$.
		\item  Irefflexive: for each $A \in \mathcal{F}$ it is not true that $A \to A$.
		\item Antisymetric: for $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
		\end{itemize}
		If $A \to B$ we will say that $A$ is \textbf{a cause} of $B$.
		\end{defi}
		\end{frame}

		\begin{frame}
		\frametitle{From Graphs to Probability Measures}
		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph (\cite{koller2009probabilistic}). \\
		We require, as axioms, that this measure satisfies:
        \begin{itemize}
        \item Causal Markov Condition.
        \item Causal Minimality.
         \item Causal Faithfulness.
         \item Causal Sufficiency
        \end{itemize}
		\end{frame}

@@

\subsection{Decision Problems}
	\begin{frame}
	Decision Problems and Rational Thinking.
	\end{frame}
	\begin{frame}
	\frametitle{Decision Problems}
	A decision problem is composed by (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}):
	\begin{itemize}
	\item A set of available options $\mathcal{A}$.
	\item A (rational) decision maker with preference relation $\succeq$ defined over $\mathcal{A} \times \mathcal{A}$.
	\item A family of uncertain events $\mathcal{E}$.
    \item A family of possible outcomes $\mathcal{C}$
	\end{itemize}
	\end{frame}

@@

\section{Research Problem}
\begin{frame}
About my Proposed Research:
\begin{enumerate}
\item Motivation.
\item Justification.
\item Research Questions.
\item Problem Statement.
\item Hypothesis.
\end{enumerate}
\end{frame}
	\subsection{Motivation}
	\begin{frame}
	\frametitle{Motivation: 1}
	\begin{itemize}
	\item Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}).
	\item Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions.
	\end{itemize}
	\end{frame}
	\begin{frame}
	\frametitle{Motivation: 2}
	\begin{itemize}
	\item It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
	\item What about \textbf{not knowing} the causal model? 
	\end{itemize}
	\end{frame}

	\subsection{Justification}
	\begin{frame}
	\frametitle{Justification: Un-interpretable models}
	\begin{itemize}
	\item Many real-world applications of decision making are solved by \textit{associative} methods.
	\item Associative methods capture only statistical patterns that are found in data.
	\item As an example, methods based in Deep Neural Networks aren't even supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
	\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Justification: Need for interpretable decision-making models}
	\begin{itemize}
	\item Current methods for Decision Making are based in Reinforcement Learning, and specifically in Deep Reinforcement Learning. 
	\item Although good performance in the task, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
	\item This is highly relevant in real-world applications such as self-driving cars.
	\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Justification: Explaining why}
	\begin{itemize}
	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be able to explain. 
	\item It allows to ask \textit{why}.
	\item Transferability
	\end{itemize}
	\end{frame}

@@

\section{Approaches and Related Work}
\begin{frame}
\frametitle{Approaches and Related Work}
Approaches and Related Work:
\end{frame}
\begin{frame}
\frametitle{Approaches and Related Work}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Related_Work.png}
\caption{Related Work}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\subsection{Approaches and Related Work: Reinforcement Learning}
\begin{frame}
\frametitle{Reinforcement Learning (RL)}
\begin{itemize}
\item Reinforcement Learning uses only associative patterns for finding an optimal policies.
\item Optimal policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
\item In this way, when only associative information is available, RL is a coherent learning framework. 
\end{itemize}
\end{frame}

\subsection{Approaches and Related Work: Causal learning in humans}
\begin{frame}
\frametitle{Causal Learning in Human beings}
\begin{itemize}
\item Human beings, while facing a sequential decision problem, use and modify causal information.
\item \cite{hagmayer2009decision}, \cite{hagmayer2013repeated} show that human beings can learn causal information in simple decision-making tasks, but remains open to study causal learning while seeking for a specific objective. 
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Joyce's Causal Decision Theory}
\begin{frame}
\frametitle{Joyce's Causal Decision Theory}
\begin{itemize}
\item An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce (\cite{joyce1999foundations}), who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes.
\item  His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Bandits}
\begin{frame}
\frametitle{Bandits: Non-causal}
\begin{itemize}
\item In the class of decision problems considered for this proposal it is assumed that the a rational agent must choose some available action and then receive a reward from it.
\item Problems of this nature have been modelled as \textit{bandit} problems.
\item Best-arm algorithms exist: none consider causal information.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Bandits: Causal bandits}
\begin{itemize}
\item An action can be conceived as an \textit{intervention} over an environment, which is in fact how humans consider their actions in the world.
\item They use and modify causal knowledge during a sequential decision making process.
\item It is shown by \cite{lattimoreNIPS2016} that adding causal information in a fixed budget decision problem allows the decision maker to learn \textit{faster} had he not considered causal information.
\item Their work requires that the causal model is fully known to the decision maker.
\item Relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part. 
\end{itemize}
\end{frame}
\section{Proposed Solution and Methodology}
\begin{frame}
\frametitle{Proposed Solution and Methodology}
Proposed Solution and Methodology:
\begin{enumerate}
\item General and Specific Objectives.
\item Methodology.
\item Validation.
\item Detailed Methodology.
\end{enumerate}
\end{frame}

	\subsection{Objectives}
	\begin{frame}
	\frametitle{General Objectives}
	The proposed research has as a general objective:
	\begin{itemize}
 \item Understand the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
	\end{itemize}
	\end{frame}

	\subsection{Specific Objectives}
	\begin{frame}
	\frametitle{Specific Objectives}
	In order to reach the General Objective we must:
	\begin{enumerate}
	\item Define a Game between agent and Nature in order to model agent-environment interaction.
	\item Find an adequate family of probability distributions that encode causal knowledge.
	\item Develop a general belief-updating criteria in terms of causal information.
	\end{enumerate}
	\end{frame}

@@

\begin{frame}
\frametitle{A conjecture}
It needs to be shown that given the rewards obtained by a decision-making procedure that achieves the MEU, then my proposed algorithm must stay close to it.
\begin{teo}
Let $(X_1,X_2,...)$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...)$ are the rewards obtained by a decision-making procedure that uses causal information in the way we propose, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$. 
\end{teo}
\begin{proof}
To be proved.
\end{proof}
\end{frame}

@@

\subsection{Proposed Solutuion: A guiding principle}
\begin{frame}
\frametitle{A guiding principle for Causal Decision Problems}
We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational.\\
We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.\\
It remains to be refined and extended for a general case. 
\end{frame}

\begin{frame}
\begin{prop}
Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent. Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
\[ P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
\end{prop}
\end{frame}

\subsection{First Step: Modelling a Causal Decision Problem as a Game}
\begin{frame}
\frametitle{First Step: Modelling a Causal Decision Problem as a Game}
\begin{itemize}
\item The next step is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
\begin{enumerate}
\item The (original) decision maker.
\item A player which will be called \textbf{Nature}, whose actions are to be guided by the causal model.
\item The game is over when the \textit{outcome} of the original decision problem is realized. This we will call a \textit{round}.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item  We assume that player Nature has the first move and she assigns some state to any, or all, of the variable(s) in the causal model $\mathcal{G}$.
\item After the decision maker makes his \textit{play}, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
\item This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games. Details in \cite{osborne1994course}.
\end{itemize}
\end{frame}

\subsection{Second Step: Belief formation}
\begin{frame}
\frametitle{Second Step: Belief Formation}
\begin{itemize}
\item The whole idea is that a decision maker is able to \textbf{learn} and \textbf{reason} about his environment by observing the effects of his actions.
\item The agent encodes any previous knowledge or uncertainties in a probability distribution.
\item The observed information will be used to modify his current knowledge in a way that is similar to how human beings modify causal knowledge while intervening in the world
\end{itemize}
\end{frame}

\subsection{Third Step: Making choices}
\begin{frame}
\frametitle{Third Step: Making Good Choices}
\begin{itemize}
\item The decision maker uses his current knowledge \textit{as if} it were the truth about the causal relations of his environment.
\item With this information he will choose the action that maximize the Expected Utility, according with our guiding principle.
\end{itemize}
\end{frame}

\subsection{Fourth Step: Belief updating}
\begin{frame}
\frametitle{Belief Updating}
\begin{itemize}
\item It is known that belief updating using Bayes' Theorem is the only way rational way of updating uncertainty.
\item It remains to be answered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way.
\end{itemize}
\end{frame}

\section{Methodology divided into simpler cases}
\begin{frame}
We will show how the methodology applies in simpler and easier cases where the decision maker has some extra info about his environment.
\end{frame}
\subsection{Fully knowing the causal graphical model}
\begin{frame}
\frametitle{Fully knowing the causal graphical model}
\begin{itemize}
\item Following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has.
\item Choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
\end{itemize}
\end{frame}

\subsection{Knowing only the graph structure}
\begin{frame}
\frametitle{Knowing only the causal graph structure}
\begin{itemize}
\item The agent encodes uncertainty over the parameters of the model as a probability distribution over a suitable space.
\item Done in such way that a \textit{local} (i.e. in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
\item The agent forms a causal model which is used as, according to the guiding principle, as if it were the true causal model.
\item Belief updating.
\end{itemize}
\end{frame}

\subsection{General case: Model unknown}
\begin{frame}
\frametitle{Model unknown}
If the model is completely unknown, two ideas appear:
\begin{enumerate}
\item If the maximum number of variables is known.
\item If it is not.
\end{enumerate}
\end{frame}
\begin{frame}
\frametitle{Idea 1 for the General Case}
\begin{itemize}
\item If the maximum number of variables is assumed to be known by the decision maker, then the situation simplifies since a decision maker could generate distributions that represent causal relations among variables.
\item For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model as in the previous case.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Idea 2 for the General Case}
\begin{itemize}
\item If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables. 
\item In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.  
\item The graph generating process must converge to the true model.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{20 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/20_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\begin{frame}
\frametitle{50 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/50_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}
